---
title: "applied machine learning analysis on real world data
date: "2026-02-27"
---
# Setting up
## Load Required Libraries
```{r}
library(arm)
library(caTools)
library(caret)
library(class)
library(cluster)
library(corrplot)
library(dendextend)
library(dplyr)
library(e1071)
library(factoextra)
library(ggcorrplot)
library(ggplot2)
library(GGally)
library(glmnet)
library(gridExtra)
library(knitr)
library(pROC)
library(randomForest)
library(rpart)
library(rpart.plot)
library(ROCR)
library(tidyverse)
```

## Load and inspect dataset for unsupervised learning
```{r}
# Set the working directory
setwd("~/Downloads/Coursework WD")

# Read dataset for unsupervised learning
data_1 <- read.csv("Country-data.csv", header = TRUE)

# Display the first few rows of the dataset
head(data_1)

# Check the structure of the dataset
str(data_1)
```

## Summary statistics
```{r}
# Summary of the data set (e.g. mean, median, min, max etc.)
summary(data_1)
```
## Check for Missing Values & Duplicates
```{r}
# Checking the missing values in each column and displaying the results
missing_values <- colSums(is.na(data_1))
print(missing_values)

# Checking for duplicate values in each column and displaying the results
duplicates <- sum(duplicated(data_1))
print(duplicates)
```
```{r}
#Remove country column from dataset 
data_numeric <- data_1[, -1]
```


## Exploration Data Analysis (EDA)

### Feature Distributions Plot
```{r}
# Convert data to long format
data1_long <- data_numeric %>%
  pivot_longer(cols = everything(), names_to = "Feature", values_to = "Value")

# Create Histogram Plot
  ggplot(data1_long, aes(x = Value)) + geom_histogram(bins = 20, fill = "deeppink", color = "black", alpha = 0.5) +
  
# Create plot title
  ggtitle("Features Distributon Plot") + 
  
# Create separate plots
  facet_wrap(~ Feature, scales = "free") + 
  
# Apply theme
  theme_classic() + 
  
# Adjust font and font size
  theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5))
```
## Correlation Matrix Heatmap 
```{r}
## Compute Correlation Matrix
cor_mat <- round(cor(data_numeric), 2)
cor_mat

## Create a Correlation Matrix Heatmap
corrplot::corrplot(cor_mat, method = 'color',
                   order = 'hclust',
                   type = "upper",
                   tl.cex = 0.8, tl.srt = 45, tl.col = "black",
                   addCoef.col = "black",
                   number.cex = 0.7, 
                   col = colorRampPalette(c("blue", "white", "#E46726"))(100))
```
### Principal Component Analysis (PCA)
```{r}
# Set a random seed for reproducibility
set.seed(123)

# Scale the data
data_scaled <- scale(data_numeric)

# Peform Principal Component Analysis
data.pca <- prcomp(data_scaled)
summary(data.pca)

# Compute the explained variance for each principal component
explained_var <- data.pca$sdev^2 / sum(data.pca$sdev^2)

# Compute the cumulative variance for each principal component
  cumulative_var <- cumsum(explained_var)

# Create a data frame for scree plot
  scree_data_frame <- data.frame(PC = 1:length(explained_var), Variance = explained_var)

# Create a scree plot
  ggplot(scree_data_frame, aes(x = PC, y = Variance)) + 

# Add points and line
  geom_point(size = 3, color = "deeppink") +
  geom_line() +

# Create plot title 
  ggtitle("Scree Plot") + 
  xlab("Principal Components") + ylab("Proportion of Variance Explained") + 

# Apply theme
  theme_classic() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 12))
```
### K means clustering

### Elbow Method
```{r}
# Set a random seed for reproducibility
set.seed(123)

# Compute Within-Cluster Sum Of Squares (wss)
wss <- sapply(1:10, function(k){
  kmeans(data_scaled, centers = k, nstart = 50, iter.max = 100)$tot.withinss
})
wss

# Create a data frame 
  elbow_df <- data.frame(k = 1:10, wss = wss)

# Plot Elbow Method Scree Plot
  ggplot(elbow_df, aes (x = k, y = wss)) +
    
# Add points and line
  geom_point(size = 3, color = "deeppink") +
  geom_line() + 
  scale_x_continuous(breaks = 1:10) + 
  
# Create plot title
  ggtitle("Elbow Method") + 
    xlab("Number of Clusters") + ylab("Within-Cluster Sum Of Squares") +
  
# Apply theme
  theme_classic() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 12))
```
```{r}
# Compute average silhouette for K CLUSTERS
sil <- sapply(2:10, function(k){
 km <- kmeans(data_scaled, centers = k, nstart = 50, iter.max = 100)
  ss <- silhouette(km$cluster, dist(data_scaled))
 mean(ss[, 3])
 })
```

### Silhouette Method 
```{r}
# Set a random seed for reproducibility
set.seed(123)
# Create a data frame for silhouette plot
sil_df <- data.frame(k = 2:10, silhouette = sil)

# Plot silhouette scores
ggplot(sil_df, aes (x = k, y = silhouette)) +
  
# Add line and points
  geom_line() + 
  geom_point(size = 3, color = "deeppink") +
  
# Highlight optimal k cluster amount
  geom_vline(xintercept = 5, color = "red") +
   scale_x_continuous(breaks = 2:10) + 
  
# Create plot title
  ggtitle("Silhouette Method") + 
  xlab("Number of Clusters K") + ylab("Average Silhouette Width") +
  
# Apply theme
  theme_classic() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 12))

```

## Apply k-means Clustering
```{r}
# Set a random seed for reproducibility
set.seed(123)

# Perform k-means clustering
kmeans_1 <- kmeans(data_scaled, centers = 3, nstart = 25)
kmeans_1

# Display the number of data points
table(kmeans_1$cluster)

# Convert to a factor
data_1$cluster <- as.factor(kmeans_1$cluster)

```

### Box plot for cluster analysis
```{r}
# Create a series of box plots
bp <- function(y_var, title){
  ggplot(data_1, aes(x = factor(cluster), y = .data[[y_var]], fill = factor(cluster))) +
    geom_boxplot() +
    
# Create plot titles and label axis
    labs(title = title, x = "Cluster", y = y_var) +
    
# Apply theme
    theme_minimal() + scale_fill_brewer(palette="Dark2") +
    theme(plot.title = element_text(hjust = 0.5), axis.title = element_text(size = 12))
    }

# Generate box plots
box_1 <- bp("child_mort","child_mort vs kclusters")
box_2 <- bp("gdpp","gdpp vs kclusters")
box_3 <- bp("income", "income vs kclusters")
box_4 <- bp("life_expec", "life_expec vs kclusters")

# Arrange box plots
grid.arrange(box_1, box_2, box_3, box_4, ncol = 2)

```
### Hierarchical Clustering
```{r}
# Calculate euclidean distance
dist_mat <- dist(data_scaled, method = 'euclidean')

# Perform hierarchical clustering using ward's method
hc <- hclust(dist_mat, method = "ward.D2")
dend <- as.dendrogram(hc)

# Plot dendrograms
dend_col <- color_branches(dend, k = 3)
plot(dend_col, 
     
# Add title and label axis
     main = "Dendrogram using Ward's Method", xlab = " ", ylab = "Height")

# Add red horizontal line to show clusters 
abline(h = 21, col = "red", lwd = 2)

```
# Box plots with clusters added

```{r}
# Cut tree to produce 3 clusters
clusters_hc <- cutree(hc, k = 3)

# Append to dataset
dataset_with_hc_clusters <- data_1

# Convert to factor
dataset_with_hc_clusters$Cluster <- as.factor(clusters_hc)

# Create a series of box plots
bp_1 <- function(y_var, title){
  ggplot(dataset_with_hc_clusters, aes(x = factor(Cluster), y = .data[[y_var]], fill = factor(Cluster))) +
    geom_boxplot() +
    
# Create plot titles and label axis
    labs(title = title, x = "Cluster", y = y_var) +
    
# Apply theme
    theme_minimal() + scale_fill_brewer(palette="Dark2") +
    
# Adjust font and font sizes
    theme(plot.title = element_text(hjust = 0.5), axis.title = element_text(size = 12))
    }

# Generate box plots
box_1 <- bp_1("child_mort", "child_mort vs kclusters")
box_2 <- bp_1("gdpp", "gdpp vs kclusters")
box_3 <- bp_1("income", "income vs kclusters")
box_4 <- bp_1("life_expec", "life_expec vs kclusters")

# Arrange box plots
grid.arrange(box_1, box_2, box_3, box_4, ncol = 2)
```
### START OF REGRESSION

# Load and inspect data for regression

```{r}
# Read dataset for regression
data_2 <- read.csv("Life Expectancy Data.csv", header = TRUE)

# Display the first few rows of the dataset
head(data_2)

# Check the structure of the dataset
str(data_2)

# Compute correlation between infant deaths and under five deaths
correlation <- cor(data_2$infant.deaths, data_2$under.five.deaths, use = "complete.obs")
print(correlation)
```
## Remove columns from dataset
```{r}
# Remove Country column from dataset
data_2 <- data_2 %>% dplyr::select(-Country)

# Remove Year column from dataset
data_2 <- data_2 %>% dplyr::select(-Year)

#Remove infant.deaths column from dataset due to high correlation with under five deaths
data_2 <- data_2 %>% dplyr::select(-infant.deaths)
```

## Check for Missing Values & Duplicates
```{r}
# Checking the missing values in each column and displaying the results
missing_values_2 <- colSums(is.na(data_2))
print(missing_values_2)

# Checking for duplicate values in each column and displaying the results
duplicates_2 <- sum(duplicated(data_2))
print(duplicates_2)
```

```{r}
# Impute the missing data
data_2 <- data_2 %>%
mutate(across(where(is.numeric),
                ~ ifelse(is.na(.),
                         median(., na.rm = TRUE),
                         .)))

# Checking for missing values after imputing
missing_values_2 <- colSums(is.na(data_2))
print(missing_values_2)
```

## Distribution
```{r}
# Create Histogram Plot
ggplot(data_2, aes(x = Life.expectancy)) + geom_histogram(aes(y = after_stat(density)), bins = 30, alpha = 0.7, fill = "deeppink", color = "pink") +
                                                            
# Add density curve
  geom_density(color = "black") +
  
# Create plot title
  labs(title = "Distribution", x = "Life Expectancy", y = "Density") +
  
# Apply theme
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 12))
```

## Correlation Matrix Heatmap 
```{r}
## Compute Correlation Matrix
cor_mat2 <- cor(data_2 %>% dplyr::select(where(is.numeric)), use = "pairwise.complete.obs")
cor_mat2

## Create a Correlation Matrix Heatmap
corrplot::corrplot(cor_mat2, method = 'color',
                   order = 'hclust',
                   type = "upper",
                   tl.cex = 0.7, tl.srt = 45, tl.col = "black",
                   addCoef.col = "black",
                   number.cex = 0.4, col = colorRampPalette(c("blue", "white", "#E46726"))(100))
```

# Linear Regression
```{r}
# Set a random seed for reproducibility
set.seed(123)

# Train / Test split in 80:20 ratio
sample_split <- sample.split(data_2$Life.expectancy, SplitRatio = 0.8)
train_set <- subset(data_2, sample_split == TRUE)
test_set  <- subset(data_2, sample_split == FALSE)

# Fit the model 
model <- lm(Life.expectancy ~ ., data = train_set)

# Obtain the summary
summary(model)

```
# Predictions and Performance on linear regression
```{r}
# Make predictions on test set
predictions <- predict(model, test_set)

# Compute the Mean Squared Error (MSE) 
mse <- mean((test_set$Life.expectancy - predictions)^2)
print(mse)

# Compute Root Mean Squared Error (RMSE)
rmse <- sqrt(mse)
print(rmse)
```

# Plot Linear Regression
```{r}

linear_results <- data.frame(Actual = test_set$Life.expectancy, Predicted = predictions)
# Create scatter plot for linear regression
ggplot(linear_results, aes(x = Actual, y = Predicted)) +

# Add line
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") + 
  geom_point(color = "deeppink") +

# Create plot title
  labs(title = "Linear: Actual vs Predicted",
    x = "Actual", y = "Predicted") +
  
# Apply theme
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 15))
```
# Polynomial Linear Regression
```{r}
# Fit a polynomial regression model
poly_model <- lm(Life.expectancy ~ poly(GDP, 2) + ., data = train_set)

# Obtain the summary
summary(poly_model)
```
# Predictions and Performance on Polynomial Regression
```{r}
# Make predictions on polynomial test set
poly_predictions <- predict(poly_model, test_set)

# Compute the Mean Squared Error (MSE)
poly_mse <- mean((test_set$Life.expectancy - poly_predictions)^2)
print(poly_mse)

# Compute Root Mean Squared Error (RMSE)
poly_rmse <- sqrt(poly_mse)
print(poly_rmse)

```
# Plot Polynomial Regression
```{r}
# Create scatter plot for polynomial regression
poly_results <- data.frame(Actual = test_set$Life.expectancy, Predicted = poly_predictions)
ggplot(poly_results, aes(x = Actual, y = Predicted)) +

geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") + 
  geom_point(color = "deeppink") +
  labs(title = "Poly: Actual vs Predicted",
    x = "Actual", y = "Predicted") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 15))
```
# Regularised Regression
## Ridge Regression
```{r}
# Set a random seed for reproducibility 
set.seed(123)

# Set up control
train_control <- trainControl(method = "repeatedcv", 
                              number = 10) # 10 fold

# Train ridge regression model with K-Fold Cross-Validation
ridge_model <- train(
  Life.expectancy ~ ., 
  data = train_set, 
  method = "glmnet", 
  tuneGrid = expand.grid(alpha = 0, lambda = seq(0.001, 1, length = 20)), 
  trControl = train_control)
print(ridge_model)

# Make predictions on ridge test set
ridge_predictions <- predict(ridge_model, test_set)

# Compute the Mean Squared Error (MSE)
ridge_mse <- mean((test_set$Life.expectancy - ridge_predictions)^2)
print(ridge_mse)

# Compute Root Mean Squared Error (RMSE)
ridge_rmse <- sqrt(ridge_mse)
print(ridge_rmse) 

```
## Plot ridge regression
```{r}
# Create scatter plot for ridge regression
ridge_results <- data.frame(Actual = test_set$Life.expectancy, Predicted = ridge_predictions)
ggplot(ridge_results, aes(x = Actual, y = Predicted)) +

geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") + 
  geom_point(color = "deeppink") +
  labs(title = "Ridge: Actual vs Predicted",
    x = "Actual", y = "Predicted") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 15))
```
## Lasso Regression 
```{r}
# Train a lasso regression model
lasso_model <- train(
  Life.expectancy ~ ., 
  data = train_set,
  method = "glmnet", 
  tuneGrid = expand.grid(alpha = 1, lambda = seq(0.001, 1, length = 20)), 
  trControl = train_control)
print(lasso_model)

# Make predictions on lasso test set
lasso_predictions <- predict(lasso_model, test_set)

# Compute the Mean Squared Error (MSE)
lasso_mse <- mean((test_set$Life.expectancy - lasso_predictions)^2)
print(lasso_mse)

# Compute Root Mean Squared Error (RMSE)
lasso_rmse <- sqrt(lasso_mse)
print(lasso_rmse) 

```

## Lasso Regression Plot
```{r}
# Create scatter plot for lasso regression
lasso_results <- data.frame(Actual = test_set$Life.expectancy, Predicted = lasso_predictions)
ggplot(lasso_results, aes(x = Actual, y = Predicted)) +

geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") + 
  geom_point(color = "deeppink") +
  labs(title = "Lasso: Actual vs Predicted",
    x = "Actual", y = "Predicted") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 15))

```
# Random forest
```{r}
# Set a random seed for reproducibility 
set.seed(123)

rf_model <- train(
  Life.expectancy ~ ., 
  data = train_set, 
  method = "rf", 
  ntree = 300,
  tuneLength = 3, 
  trControl = train_control, 
  importance = TRUE)

print(rf_model)

# Make predictions on random forest test set
rf_predictions <- predict(rf_model, test_set)

# Compute the Mean Squared Error (MSE)
rf_mse <- mean((test_set$Life.expectancy - rf_predictions)^2)
print(rf_mse)

# Compute Root Mean Squared Error (RMSE)
rf_rmse <- sqrt(rf_mse)
print(rf_rmse) 

```
# Random Forest Plot
```{r}
# Create scatter plot for Random F  orest 
rf_results <- data.frame(Actual = test_set$Life.expectancy, Predicted = rf_predictions)
ggplot(rf_results, aes(x = Actual, y = Predicted)) +

geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") + 
  geom_point(color = "deeppink") +
  labs(title = "Random Forest: Actual vs Predicted",
    x = "Actual", y = "Predicted") +
  theme_minimal() +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 15))

```

# Compare all models
```{r}
model_results <- data.frame(
  Model = c("Linear", "Polynomial", "Ridge", "Lasso", "Random Forest"), 
  RMSE = c(rmse, poly_rmse, ridge_rmse, lasso_rmse, rf_rmse))

print(model_results)

```
# bar plot to compare
```{r}
ggplot(model_results, aes(x = Model, y = RMSE, fill = Model)) +
  geom_col(fill = "deeppink", alpha = 0.7) + 
  coord_flip() +
  theme_minimal() +
  labs(title = "Comparison of Models", y = "RMSE", x = "Model") +
  theme(plot.title = element_text(size = 20, hjust = 0.5), axis.title = element_text(size = 15))
  
```

# Feature Importance
```{r}
# Set a seed for reproducibility
set.seed(123)

# Train a final Random Forest model for feature importance
importance_rf <- varImp(rf_model)
print(importance_rf)
plot(importance_rf, top = 10)
```
### Start of Classification
```{r}
# Load and inspect data for classification
# Read dataset for classification
data_3 <- read.csv("diabetes_binary.csv", header = TRUE)

# Display the first few rows of the dataset
head(data_3)

# Check the structure of the dataset
str(data_3)

 # add table fn here *** oi (^_^)
```

```{r}
## Check for Missing Values & Duplicates
# Checking the missing values in each column and displaying the results
missing_values3 <- colSums(is.na(data_3))
print(missing_values3)

# Checking for duplicate values in each column and displaying the results
duplicates_3 <- sum(duplicated(data_3))
print(duplicates_3)
```
# Remove duplicated values
```{r}
# Remove duplicated data
data_3 <- data_3[!duplicated(data_3), ]

# Checking for duplicate values
duplicates_3 <- sum(duplicated(data_3))
print(duplicates_3)

table(data_3$class)
prop.table(table(data_3$class))

```

# Bar chart to show distribution
```{r}
# Create a bar chart 
ggplot(data_3, aes(x = factor(Diabetes_binary), fill = factor(Diabetes_binary))) +
  geom_bar() +
  labs(title = "Distribution of Diabetes",
       x = "0 = No Diabetes, 1 = Diabetes",
       y = "Count",
       fill = "Diabetes") +
  theme_minimal()


```
# Correlation heatmap matrix
```{r}
## Compute Correlation Matrix
cor_mat3 <- cor(data_3 %>% dplyr::select(where(is.numeric)), use = "pairwise.complete.obs")
cor_mat3

## Create a Correlation Matrix Heatmap
corrplot::corrplot(cor_mat3, method = 'color',
                   order = 'hclust',
        
                   tl.cex = 0.7, tl.col = "black",
                   addCoef.col = "black",
                   number.cex = 0.4, col = colorRampPalette(c("blue", "white", "#E46726"))(100))

```

# Splitting training and testing sets
```{r}
# Set seed for reproducibility 
set.seed(123)

# Split data into 80% training and 20% testing
train_index <- createDataPartition(data_3$Diabetes_binary, p = 0.8, list = FALSE)

# Create training data set
train_data <- data_3[train_index, ]

# Create testing data set 
test_data <- data_3[-train_index, ]

train_data$Diabetes_binary <- factor(train_data$Diabetes_binary,
                                     levels = c(0,1),
                                     labels = c("No", "Yes"))

test_data$Diabetes_binary <- factor(test_data$Diabetes_binary,
                                     levels = c(0,1),
                                     labels = c("No", "Yes"))

# Check class distribution 
table(train_data$Diabetes_binary)
table(test_data$Diabetes_binary)
```

# Logistic Regression
```{r}
  # Train a logistic regression model
 logit_model <- glm(Diabetes_binary ~ ., data = train_data, family = binomial)
  
 logit_prob <- predict(logit_model, newdata = test_data, type = "response")
 logit_pred <- ifelse(logit_prob > 0.5, "Yes", "No")
 
 logit_pred <- factor(logit_pred)
 log_actual <- factor(test_data$Diabetes_binary)
 
# Create and print the confusion matrix for logistic regression
conf_mat_logit <- confusionMatrix(logit_pred, log_actual, positive = "Yes")
print(conf_mat_logit)

 # Compute ROC
logit_roc <- roc(test_data$Diabetes_binary, logit_prob)

```

# Random forest
```{r}
# Set up control
train_data$Diabetes_binary <- factor(train_data$Diabetes_binary)
test_data$Diabetes_binary <- factor(test_data$Diabetes_binary)
```

# Training a Random Forest (RF) model
```{r}
# Train a random forest model
rf_model <- randomForest(Diabetes_binary ~., data = train_data, ntree = 300)

rf_pred <- predict(rf_model, newdata = test_data)

rf_prob <- predict(rf_model, newdata = test_data, type = "prob")[, "Yes"]

# Create and print the confusion matrix for RF
conf_mat_rf <- confusionMatrix(rf_pred, test_data$Diabetes_binary, positive = "Yes")
print(conf_mat_rf)

# Compute ROC
rf_roc <- roc(test_data$Diabetes_binary, rf_prob)
```

# K-Nearest Neighbour (K-NN)
```{r}
# Scaling the results
preprocess <- preProcess(train_data[, -which(names(train_data) == "Diabetes_binary")],
                        method = c("center", "scale"))

train_data_scaled <- predict(preprocess, train_data)
test_data_scaled  <- predict(preprocess, test_data)

```

```{r}
# Set seed for reproducibility
set.seed(123)

# Train a K-NN model 
knn_pred <- knn(
  train = train_data_scaled[, -which(names(train_data_scaled) == "Diabetes_binary")],
  test = test_data_scaled[, -which(names(test_data_scaled) == "Diabetes_binary")],
  cl = train_data_scaled$Diabetes_binary,
  k = 5,
  prob = TRUE 
)

# Create and print a confusion matrix for K-NN
conf_mat_knn <- confusionMatrix(
  knn_pred,
  test_data_scaled$Diabetes_binary,
  positive = "Yes"
)


knn_prob <- attr(knn_pred, "prob")
knn_prob <- ifelse(knn_pred == "Yes", knn_prob, 1 - knn_prob)

# Compute ROC
knn_roc <- roc(test_data_scaled$Diabetes_binary, knn_prob)

```

# Support Vector Machine (SVM) 
```{r} 
# Set seed for reproducibility
set.seed(123)

# Sample data set
train_small <- train_data[sample(1:nrow(train_data), 5000),]

# Train a SVM model
svm_model <- svm(
  Diabetes_binary ~ .,
  data = train_small,
  kernel = "radial",
  probability = TRUE
)

# Predict 
svm_pred <- predict(svm_model, newdata = test_data, probability = TRUE)

# Probabilities
svm_prob_attr <- attr(svm_pred, "probabilities")
svm_prob <- svm_prob_attr[, "Yes"]

svm_pred <- factor(svm_pred, levels = levels(test_data$Diabetes_binary))

# Create and print the confusion matrix for SVM
conf_mat_svm <- confusionMatrix(
  svm_pred,
  test_data$Diabetes_binary,
  positive = "Yes"
)
print(conf_mat_svm)

 # Compute ROC
svm_roc <- roc(test_data$Diabetes_binary, svm_prob)

```

# Plot curves
```{r}
# Plot ROC curves for all models

plot(logit_roc, col = "blue", main = "ROC Curves for All Models", legacy.axes = TRUE, xaxs = "i", lwd = 2)

plot(rf_roc, col = "deeppink", add = TRUE, lwd = 2)

plot(knn_roc, col = "orange", add = TRUE, lwd = 2)

plot(svm_roc, col = "purple", add = TRUE, lwd = 2)

legend("bottomright",
       legend = c("Logistic Regression", "Random Forest", "K-NN", "SVM"),
       col = c("blue", "deeppink", "orange", "purple"),
       lwd = 2)

```

# Compute Area Under Curve (AUC) and F1
```{r}
# Collect metrics
model_metrics <- data.frame(
  Model = c("Logistic Regression", "Random Forest", "KNN", "SVM"),

  Accuracy = c(
    confusionMatrix(logit_pred, test_data$Diabetes_binary)$overall["Accuracy"],
    confusionMatrix(rf_pred, test_data$Diabetes_binary)$overall["Accuracy"],
    conf_mat_knn$overall["Accuracy"],
    conf_mat_svm$overall["Accuracy"]
  ),

  F1_Score = c(
    confusionMatrix(logit_pred, test_data$Diabetes_binary)$byClass["F1"],
    confusionMatrix(rf_pred, test_data$Diabetes_binary)$byClass["F1"],
    conf_mat_knn$byClass["F1"],
    conf_mat_svm$byClass["F1"]
  ),

  ROC_AUC = c(
    auc(logit_roc),
    auc(rf_roc),
    auc(knn_roc),
    auc(svm_roc)
  )
)

# View results
print(model_metrics)

```

```{r}
# Plot Accuracy
ggplot(model_metrics, aes(x = Model, y = Accuracy, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(title = "Accuracy Comparison", y = "Accuracy") +
  theme_minimal()

# Plot F1 Score
ggplot(model_metrics, aes(x = Model, y = F1_Score, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(title = "F1 Score Comparison", y = "F1 Score") +
  theme_minimal()

# Plot ROC auc
ggplot(model_metrics, aes(x = Model, y = ROC_AUC, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(title = "ROC AUC Comparison", y = "AUC") + 
  theme_minimal()
```





